\large How resource abundance and stochasticity affect organisms' space-use requirements

### 2024-03-06 writing/manuscript.Rmd
moved stuff to appendices; working on discussion

\noindent \textbf{Figures}: 5

@@

* make sure appendices flow (added stuff from main text)

@@

\noindent where $H$ is the organism's home-range size, $C$ is the organism's resource consumption (kcal day$^{-1}$), and $R$ is the resources the organism can access (kcal day$^{-1}$ unit area$^{-1}$). Harestad and Bunnel's model is simple to conceptualize and allows for testable predictions, but it fails to account for the nonlinear relationship required to avoid negative home range estimates [e.g., Figure 2 in the work of @bista_effect_2022].<!-- to an organism's values of $C$ and $R$ likely depend on numerous factors, such as competition, metabolic rate, diet, and body weight [@harvey_primate_1981; @lindstedt_home_1986; @reiss_scaling_1988; @harvey_primate_1981; @gittleman_carnivore_1982; @jetz_scaling_2004; @boratynski_energetic_2020; @noonan_effects_2020].--> Many researchers have since demonstrated that organisms adapt their home ranges in response to resources abundance, but studies often fail to build upon previous work. Instead, results are typically reported as independent, novel findings. Thus, there remains a need for a clear hypothesis for the effects of resource abundance on organisms' space-use requirements. Additionally, while much work has been done on estimating organisms' responses to average resource abundance, there is little information on how they respond to unpredictable changes in resources [but see: @stephens_optimal_1982; @rizzuto_forage_2021].

@@

\noindent While organisms' needs vary greatly between taxonomic groups, some needs are essential for most species for survival and reproduction. All heterotrophic organisms require sources of chemical energy (i.e., food), water, and various limiting nutrients to survive, grow, and reproduce [@harvey_primate_1981; @baldwin_nutritional_1984; @reich_body_2001]. Failing to acquire sufficient resources can result in lower fitness, behavioral changes, physiological changes or damage, inability to reproduce, and death. Motile organisms can move to new locations in search of resources [@charnov_optimal_1976; @brown_ecology_1999], but they must also account for a higher metabolism and movement costs [@taylor_energetics_1982]. The hypothesis that space-use requirements decrease with resource abundance, $\e(R)$, is commonly accepted and well supported, but many studies assume a linear relationship [e.g., @harestad_home_1979; @relyea_home_2000; @rizzuto_forage_2021; @bista_effect_2022; @bradsworth_using_2022; @mcclintic_effects_2014]. This is problematic because a linear relationship can lead to questionable estimates and problematic biases, such as negative or excessively small home ranges [e.g., figure 2 in the work of @bista_effect_2022]. Conceptually, the relationship between space-use requirements and $\e(R)$ must be nonlinear, since there are limits to how much space an organism is able to explore in its lifetime and a minimum amount of space it requires to survive [which is necessarily greater than zero -- see: @lucherini_habitat_1996; @herfindal_prey_2005; @nilsen_can_2005; @simcharoen_female_2014; @watson_ferruginous_2020, and contrast them to the estimates based on linear models listed above].

@@

Consider the scenario where $\e(R) = \mu(t)$ changes over time but $\var(R) = \sigma^2 > 0$ is constant over time and space (Fig. A1).<!-- We can use the notation $R \sim \Gamma\big(\mu(t), \sigma^2\big)$ to indicate that $R$ follows a Gamma distribution with a mean $\mu(t)$ that changes over time (i.e., it is a function of time) while the variance is constant.--> Since $R$ is random, an organism that moves in the landscape will not find the same $R$ at different time points or throughout the habitat. However, if $\mu(t)$ changes repetitively and regularly over time (e.g. peaks each spring and is lowest in winter), an organism may learn to predict times of high or low $\e(R)$ [e.g., @falcon-cortes_hierarchical_2021; @geremia_migrating_2019; @abrahms_memory_2019; @samarra_movements_2017] but will still not be able to predict $R$ exactly.

@@

***HERE***

@@

\noindent We have provided the case for why both $\e(R)$ and $\var(R)$ affect organisms' space-use requirements, but we presented the two parameters as independent drivers of movement. However, in the more realistic scenario where both $\e(R)$ and $\var(R)$ fluctuate over time, organisms may respond to changes in $\var(R)$ more when resources are scarce than when they are abundant. Consequently, models estimating the effects of $R$ on organism's movement should account for not only the effects of $\e(R)$ and $\var(R)$ separately, but also their interactive effects. A highly unpredictable habitat may be very inhospitable if resources are poor, but $\var(R)$ may have little effect if resources are stochastic but always abundant. Thus, we expect $\var(R)$ to have a stronger effect on space-use requirements when $\e(R)$ is low, and less of an effect when $\e(R)$ is high.

@@

## Simulating responses to $\e(R)$ and $\var(R)$

\noindent To support our hypotheses of how organisms' space use is affected by $\e(R)$, $\var(R)$, and the interaction effect of $\e(R)$ and $\var(R)$, we present the results from a series of quantitative simulations. To start, we used the `ctmm` package [@fleming_ctmm_2021] for `R` [@r_core_team_r_2023] to generate 200 tracks (see Appendices 2 for sensitivity analyses) from an Integrated Ornstein-Uhlenbeck movement model [IOU model, see @gurarie_correlated_2017]. The IOU model's correlated velocity produces realistic tracks with directional persistence despite the tracks being discrete samples of continuous-time movement processes. However, unlike Ornstein-Uhlenbeck (OU) and Ornstein-Uhlenbeck Foraging (OUF) models, IOU models do not assume the movement process is spatially stationary, so the organism is not assumed to be range-resident. Consequently, each track is spatially unrestricted and can be interpreted as purely exploratory movement.

Each of the 200 tracks were placed on a raster with common starting point $\langle 0, 0\rangle$ and sufficient time between tracks to be independent of one another (other than the starting point). Each time the track moved to a new cell, the organism collected resources $R$ sampled from a Gamma distribution. The mean and variance of the distribution were defined by deterministic functions $\mu(t)$ and $\sigma^2(t)$ (orange and blue lines in fig. \@ref(fig:5-5-sims)). Note that $t$ was constant within each set of 200 tracks, so the time spent moving by the organism in each track did not affect the distribution $R$ was sampled from. Tracks were truncated once the organism reached satiety, and the organism was given enough time to return to $\langle 0, 0\rangle$ with no correlation to the following track. Finally, each set of 200 truncated tracks was modeled using an OUF model and Autocorrelated Kernel Density Estimate to estimate the 95% utilization distribution. Additional information is provided in Appendix B.

@@

Fig. \@ref(fig:5-5-sims) shows how the simulated space-use requirements changed in response to $\mu(t)$ and $\sigma^2(t)$. The top row (constant $\var(R)$) shows how space-use requirements vary for different trends in $\mu(t)$ while $\var(R)$ remains constant (like in fig. A1). As $\e(R)$ increases at a constant slope (linear $\mu(t)$) the space-use requirements decrease nonlinearly, with larger changes when $\e(R)$ is low, until home-range size approaches the minimum size required by the organism. As described in the section on the effects of resource abundance, the nonlinear decrease is because changes in $\mu(t)$ have a larger effect when $\e(R)$ is low [e.g., @lucherini_habitat_1996; @nilsen_can_2005; @herfindal_prey_2005].

***HERE***

In regions where the $\mu(t)$ changes over time [e.g., seasonal changes -- see cyclical $\e(R)$ and @lai_movement_2017], organisms should have access to sufficient space to fulfill their needs during periods of scarcity, whether the space available is sufficiently large year-round or it changes seasonally with the expected changes in $\mu(t)$ (e.g., winter park closures). However, estimates of spatial requirements based on estimated changes in $\mu(t)$ should be interpreted carefully, since model error and unforeseeable decreases in $\e(R)$ (such as following fires or floods) may increase organisms' space-use requirements suddenly and unpredictably. Thus, it is best to include a "buffer" area so the available space is larger than the estimated space-use requirements. This is particularly the case in environments where resource abundance changes unpredictably (drifting $\mu(t)$), since accurate long-term estimates and predictions of $\mu(t)$ may be hard to produce, if not impossible. In cases where $\mu(t)$ is highly unpredictable, organisms should have enough space to collect resources during times of greatest predicted scarcity, particularly if the changes in resource abundance occur rapidly, often, or for long periods of time (erratic $\mu(t)$).

The leftmost column of fig. \@ref(fig:5-5-sims) (constant $\e(R)$) illustrates the effects of $\var(R)$ on the simulated space-use requirements while $\e(R)$ remains constant. Overall, both mean $H$ and the variance around it increase with $\sigma^2(t)$ (most visible with constant $\e(R)$ and linear $\var(R)$). This is because, similarly to resource-poor periods, times of greater stochasticity require the organism to move over larger areas for longer periods of time. This results in a greater uncertainty in how much time and space the organism will require to reach satiety, or indeed whether an organism living in highly stochastic environments can even reach satiety within a finite amount of time.

The remaining panels in fig. \@ref(fig:5-5-sims) illustrate how the effect of $\var(R)$ depends on $\e(R)$ (and vice-versa) as well as how complex the relationship can be. Since $\e(R)$ and $\var(R)$ have opposite effects on $H$, disentangling the effects can be particularly difficult when both parameters change in a correlated manner (e.g., linear $\e(R)$ and $\var(R)$). When both $\e(R)$ and $\var(R)$ increase linearly, $H$ initially increases since the effect of $\var(R)$ is stronger, but then decreases as the effect of $\e(R)$ begins to dominate. However, when $\mu(t)$ and $\sigma^2(t)$ follow fairly different trends (e.g., cyclical $\e(R)$ and linear $\var(R)$), it is easy to see how an increase in $\var(R)$ amplifies the effects of $\e(R)$. In contrast, the effect of $\var(R)$ is stronger when $\e(R)$ is low (e.g., linear $\e(R)$ with cyclical $\var(R)$). Thus, $H$ is largest when $\e(R)$ is low and $\var(R)$ is high, and it is smallest when $\e(R)$ is high and $\var(R)$ is low [see the panel with cyclical $\e(R)$ and $\var(R)$ as well as @lai_movement_2017].

Not all 25 scenarios depicted in fig. \@ref(fig:5-5-sims) may be realistic, but the trends in $\e(R)$ and $\var(R)$, and their impacts on space use are useful examples that can be thought of as simplified scenarios. $\e(R)$ and $\var(R)$ may be assumed to be (approximately) constant in highly homogeneous environments. Although it is impossible for $\mu(t)$ and $\sigma^2(t)$ to increase linearly forever, these examples are useful to demonstrate that linear changes in $\mu(t)$ and $\sigma^2(t)$ affect $H$ nonlinearly. Cyclical oscillations in $\e(R)$ and $\var(R)$ may occur in urban environments as human activity changes within and between days [@peron_periodic_2017; @ikeda_home_2022] and as temperatures fluctuate daily and seasonally [@geremia_migrating_2019; @alston_temperature_2020], while $\e(R)$ and $\var(R)$ may drift randomly in highly complex environments which are too hard to predict. Finally, erratic changes in $\e(R)$ and $\var(R)$ may occur in environments where changes are very sudden, such as areas prone to fires or floods, as well as habitats with drastic human alteration (e.g., a forest which is clear-cut with a subsequent artificial re-forestation). However, if stochastic changes are sufficiently small and frequent, organisms may perceive them as continuous and smooth changes rather a series of small and sudden changes.

Although the temporal trends in fig. \@ref(fig:5-5-sims) are complex and the effects of $\e(R)$ and $\var(R)$ can be hard to disentangle, two simple relationships emerge when home-range size is shown as a function of either $\e(R)$ or $\var(R)$, rather than time. These relationships follow the hypotheses we presented in figs. \@ref(fig:mu-hyp) and \@ref(fig:s2-hyp) and are represented in panels A and B of fig. \@ref(fig:5-5-reg), while panel C shows the amount by which the effect of $\mu(t)$ on $H$ changes as a function of $\sigma^2(t)$ (and vice-versa; panel C).<!--, although we found that the effect of $\var(R)$ is closer linear than we expected (it is mostly sub-linear as long as $H$ is sufficiently far from zero). We ensured this was not due to a low maximum $\var(R)$ (as in the left-hand side of fig. \@ref(fig:s2-hyp)) by simulating within the range $\sigma^2(t) \in [2\,500, 250\,000] \implies \sigma \in [50, 500]$, since the range of $\e(R)$ was $\mu(t) \in [100, 200]$.--> We believe the approximately linear effect of $\var(R)$ on $H$ is mainly due to two factors. Firstly, $H$ scales superlinearly with the standard deviation in $R$ (as $\sqrt{\var(R)}$ is on the same scale as $\e(R)$; not shown), which would necessarily imply that $H$ increases sublinearly with $\var(R)$. Secondly, the organism's movement was simulated to be range resident within an homogeneous environment. Since real-world landscapes are spatially heterogeneous in both $\e(R)$ and $\var(R)$, organisms may choose to shift their home range to a new area following a stochastic event that caused $\mu(t)$ to decrease, such as a fire or a flood. Thus, we hypothesize that our simulations may be underestimating organisms' responses to spatiotemporal changes in $\e(R)$ and $\var(R)$, since the simulated organisms cannot select for areas of higher $\e(R)$ or lower $\var(R)$. Additionally, the simulations also ignore how competition, movement costs, predation, and other factors may affect the effects of $\e(R)$ and $\var(R)$ on $H$. While these are definitely important effects that should be accounted for [@jetz_scaling_2004], doing so was beyond the scope of this paper.

```{r 5-5-reg, eval=TRUE, echo=FALSE, fig.cap="Effects of $\\e(R)$ and $\\var(R)$ on simulated spatial requirements with 95\\% credible intervals for the mean (overlapping the line). Credible intervals were calculated assuming a gaussian posterior distribution on the link scale. The relationships were estimated using a Generalized Additive Model for Location and Scale (GAMLS; Rigby and Stasinopoulos 2005; Stasinopoulos and Rigby 2007) with a Gamma location-scale family of distributions ($\\tt{mgcv\\!::\\!gammals})$. The model accounted for the effects of $\\e(R)$ and $\\var(R)$ and the interaction effects between the two on both the location (i.e., mean) of $H$ and its scale parameter. Note the nonlinear decrease in $H$ as $\\e(R)$ increases and the increase in $H$ as $\\var(R)$ increases. Additionally, note how the variance in space-use requirements strongly increases with mean space-use requirements.", out.width='100%'}
knitr::include_graphics('../figures/simulation-regression-plots.png')
```

***to do***:

* add a paragraph and figure on effects of $\mu(t)$ and $\sigma^2(t)$ on $\var(R)$.

# A lowland tapir in the Brazilian Cerrado: a case study

***to do***:

* open with limitations of sims
* then highlight shortcomings of empirical data
* move trasnformation stuff to the appendix
* move modeling details to appendix C

\noindent In this section, we show how this framework can be applied to better understand the movement ecology of free-ranging organisms via remote sensing data such as NDVI [@pettorelli_normalized_2011]. To illustrate the methods, we use empirical tracking data on a lowland tapir from the Brazilian Cerrado. Appendix C contains additional information on how NDVI was modeled along with details on the continuous-time movement models [@fleming_ctmm_2021; @noonan_scale-insensitive_2019] and autocorrelated kernel density estimation [@noonan_comprehensive_2019; @alston_mitigating_2022; @silva_autocorrelationinformed_2022] used to quantify the tapir's space-use requirements.

The hypotheses we present here allow researchers to combine otherwise complicated findings (e.g., fig. \@ref(fig:5-5-sims)) and generalize results to a single, common set of functions. We expect the two hypotheses we present here (figs. \@ref(fig:mu-hyp) and \@ref(fig:s2-hyp)) to be applicable to all motile organisms, once differences in size, metabolic needs, and diet are accounted for. In this subsection, we illustrate how researchers can leverage these hypotheses using their own data to build on previous work rather than listing the results as new findings. Additionally, the analyses can be expanded to the population or species level with the use of Hierarchical Generalized Additive Models (HGAMs) with individual-level intercepts and smooth terms [see models "GS" and "GI" in @pedersen_hierarchical_2019]. Such models would allow one to test whether members of the same population or species respond similarly to $\e(R)$ and $\var(R)$ while also quantifying any deviation from the hypothesized behavior. Individuals and populations that were once thought to have different space-use requirements due to differences in behavior may instead be simply responding to different local conditions along the functions in figs. \@ref(fig:mu-hyp) and \@ref(fig:s2-hyp) [e.g., @singh_migration_2012]. In the following sections, we illustrate how one can apply this framework to empirical data.

We start by providing examples and considerations on how one may model $R$ and the effects of $\e(R)$ and $\var(R)$ on organisms' space-use requirements. Next, we apply the methods to the tapir's GPS tracking data and use NDVI as a proxy for $R$. Finally, we offer suggestions on how this approach can be used to inform conservation-related decisions, including assessing habitat quality and estimating organisms' space-use requirements under different scenarios.

### Modeling $R$ {#modeling-r}

\noindent Location-scale models [@stasinopoulos_generalized_2007; @rigby_generalized_2005] are a class of statistical models that allow us to estimate changes in a random variable's mean (i.e. its location) and variance (which depends on its scale) while allowing the mean-variance relationship to vary. `mgcv` [@wood_generalized_2017] is a commonly used package for `R` [@r_core_team_r_2023] that allows one to fit Generalized Linear Models [GLMs, see @zuur_mixed_2009] and Generalized Additive Models [GAMs, see @wood_generalized_2017], including hierarchical and location-scale GLMs and GAMs. Currently, the `mgcv` package allows one to fit location-scale models with various families of distributions, including Gaussian (i.e., normal), gamma, and Tweedie location-scale families.

The Gaussian location-scale family of distributions is very flexible, since the mean and variance parameters are assumed to be independent, and the response can be either positive or negative. However, the distribution's flexibility can also result in unreliable estimates for non-Gaussian responses, such as strictly positive data (e.g. available biomass), count data (e.g., number of prey), proportions (e.g., percentage of forested habitat), and bounded ratios [e.g., NDVI, see @pettorelli_normalized_2011].

>>>>

\noindent \textbf{Figures}: 6

@@

* make sure appendices flow (added stuff from main text)
* check number of references

@@

\noindent where $H$ is the organism's home-range size, $C$ is the organism's resource consumption (kcal day$^{-1}$), and $R$ is the resources the organism can access (kcal day$^{-1}$ unit area$^{-1}$). Harestad and Bunnel's model is simple to conceptualize and allows for testable predictions, but it fails to account for the nonlinear relationship required to avoid negative home range estimates [e.g., fig. 2 in the work of @bista_effect_2022].<!-- to an organism's values of $C$ and $R$ likely depend on numerous factors, such as competition, metabolic rate, diet, and body weight [@harvey_primate_1981; @lindstedt_home_1986; @reiss_scaling_1988; @harvey_primate_1981; @gittleman_carnivore_1982; @jetz_scaling_2004; @boratynski_energetic_2020; @noonan_effects_2020].--> Many researchers have since demonstrated that organisms adapt their home ranges in response to resources abundance, but studies often fail to build upon previous work. Instead, results are typically reported as independent, novel findings. Thus, there remains a need for a clear hypothesis for the effects of resource abundance on organisms' space-use requirements. Additionally, while much work has been done on estimating organisms' responses to average resource abundance, there is little information on how they respond to unpredictable changes in resources [but see: @stephens_optimal_1982; @rizzuto_forage_2021].

@@

\noindent While organisms' needs vary greatly between taxonomic groups, some needs are essential for most species for survival and reproduction. All heterotrophic organisms require sources of chemical energy (i.e., food), water, and various limiting nutrients to survive, grow, and reproduce [@harvey_primate_1981; @baldwin_nutritional_1984; @reich_body_2001]. Failing to acquire sufficient resources can result in lower fitness, behavioral changes, physiological changes or damage, inability to reproduce, and death. Motile organisms can move to new locations in search of resources [@charnov_optimal_1976; @brown_ecology_1999], but they must also account for a higher metabolism and movement costs [@taylor_energetics_1982]. The hypothesis that space-use requirements decrease with resource abundance, $\e(R)$, is commonly accepted and well supported, but many studies assume a linear relationship [e.g., @harestad_home_1979; @relyea_home_2000; @rizzuto_forage_2021; @bista_effect_2022; @bradsworth_using_2022; @mcclintic_effects_2014]. This is problematic because a linear relationship can lead to questionable estimates and problematic biases, such as negative or excessively small home ranges [e.g., fig. 2 in the work of @bista_effect_2022]. Conceptually, the relationship between space-use requirements and $\e(R)$ must be nonlinear, since there are limits to how much space an organism is able to explore in its lifetime and a minimum amount of space it requires to survive [which is necessarily greater than zero -- see: @lucherini_habitat_1996; @herfindal_prey_2005; @nilsen_can_2005; @simcharoen_female_2014; @watson_ferruginous_2020, and contrast them to the estimates based on linear models listed above].

@@

Consider the scenario where $\e(R) = \mu(t)$ changes over time but $\var(R) = \sigma^2 > 0$ is constant over time and space (fig. A1).<!-- We can use the notation $R \sim \Gamma\big(\mu(t), \sigma^2\big)$ to indicate that $R$ follows a Gamma distribution with a mean $\mu(t)$ that changes over time (i.e., it is a function of time) while the variance is constant.--> Since $R$ is random, an organism that moves in the landscape will not find the same $R$ at different time points or throughout the habitat. However, if $\mu(t)$ changes repetitively and regularly over time (e.g. peaks each spring and is lowest in winter), an organism may learn to predict times of high or low $\e(R)$ [e.g., @falcon-cortes_hierarchical_2021; @geremia_migrating_2019; @abrahms_memory_2019; @samarra_movements_2017] but will still not be able to predict $R$ exactly.

@@

@@

\noindent We have provided the case for why both $\e(R)$ and $\var(R)$ affect organisms' space-use requirements, but we presented the two parameters as independent drivers of movement. However, organisms may respond to changes in $\var(R)$ more when resources are scarce than when they are abundant. Consequently, models estimating the effects of $R$ on organism's movement should account for not only the marginal effects of $\e(R)$ and $\var(R)$ but also their interactive effects. A highly unpredictable habitat may be very inhospitable if resources are poor, but $\var(R)$ may have little effect if resources are stochastic but always abundant. Thus, we expect $\var(R)$ to have a stronger effect on space-use requirements when $\e(R)$ is low, and less of an effect when $\e(R)$ is high. We explore this more in the following section.

@@

## Simulating responses to $\e(R)$ and $\var(R)$ {#sims}

\noindent To support our hypotheses of how organisms' space use is affected by $\e(R)$, $\var(R)$, and the interaction effect of $\e(R)$ and $\var(R)$, we present the results from a series of quantitative simulations. To start, we used the `ctmm` package [@fleming_ctmm_2021] for `R` [@r_core_team_r_2023] to generate 200 tracks (see Appendix B for sensitivity analyses) from an Integrated Ornstein-Uhlenbeck movement model [IOU model, see @gurarie_correlated_2017]. The IOU model's correlated velocity produces realistic tracks with directional persistence despite the tracks being discrete samples of continuous-time movement processes. However, unlike Ornstein-Uhlenbeck (OU) and Ornstein-Uhlenbeck Foraging (OUF) models, IOU models do not produce spatially stationary movement, so the organism is not to range-resident. Consequently, each track is spatially unrestricted and can be interpreted as purely exploratory movement.

Each of the 200 tracks were placed on a raster with common starting point $\langle 0, 0\rangle$ and sufficient time between tracks to be independent of one another (other than the starting point). Each time the track moved to a new cell, the organism collected resources $R$ sampled from a Gamma distribution. The mean and variance of the distribution were defined by deterministic functions $\mu(t)$ and $\sigma^2(t)$ (orange and blue lines in fig. \@ref(fig:5-5-sims)). The value of $t$ was constant within each set of 200 tracks, so the time spent moving by the organism in each track did not affect the distribution $R$ was sampled from. Tracks were truncated once the organism reached satiety, and the organism was given enough time to return to $\langle 0, 0\rangle$ independently from the following track. Finally, as each set of 200 truncated was now spatially restricted and centered around $\langle 0, 0\rangle$, the set of tracks was modeled to calculate the 95% utilization distribution from an OUF model and Autocorrelated Kernel Density Estimate. We designed the simulations presented here to estimate the effects of $\e(R)$ and $\var(R)$ in simplistic environments. Consequently, we made the following four assumptions: (1) $\e(R)$ and $\var(R)$ are constant over space and short periods of time (i.e., within each set of 200 tracks, but not over $t$). (2) Organisms move randomly in the environment with no energetic costs. (3) Organisms have no perceptive range or memory, so they are unable to detect, learn, or predict where resources are abundant (high $\e(R)$) or reliable (low $\var(R)$) over time. (4) Resources do not deplete, and there is no competition or predator avoidance. Additional information is provided in Appendix B.

@@

Fig. \@ref(fig:5-5-sims) shows how simulated space-use requirements responded to changes in $\mu(t)$ and $\sigma^2(t)$ in scenarios where both functions can remain constant, increase linearly, oscillate cyclically, drift stochastically, or change erratically. The top row <!--(constant $\var(R)$) -->shows how space-use requirements vary for different trends in $\mu(t)$ while $\var(R)$ remains constant (like in fig. A1). As $\e(R)$ increases at a constant slope (linear $\mu(t)$) $H$ decreases nonlinearly, with larger changes when $\e(R)$ is low, until it approaches the minimum size required by the organism. Although less evident, note how the noise in the lines also decreases as $\e(R)$ increases.<!-- As described in the section on the effects of resource abundance, the nonlinear decrease is because changes in $\mu(t)$ have a larger effect when $\e(R)$ is low [e.g., @lucherini_habitat_1996; @nilsen_can_2005; @herfindal_prey_2005].-->

<!-- In regions where the $\mu(t)$ changes over time [e.g., seasonal changes -- see cyclical $\e(R)$ and @lai_movement_2017], organisms should have access to sufficient space to fulfill their needs during periods of scarcity, whether the space available is sufficiently large year-round or it changes seasonally with the expected changes in $\mu(t)$ (e.g., winter park closures). However, estimates of spatial requirements based on estimated changes in $\mu(t)$ should be interpreted carefully, since model error and unforeseeable decreases in $\e(R)$ (such as following fires or floods) may increase organisms' space-use requirements suddenly and unpredictably. Thus, it is best to include a "buffer" area so the available space is larger than the estimated space-use requirements. This is particularly the case in environments where resource abundance changes unpredictably (drifting $\mu(t)$), since accurate long-term estimates and predictions of $\mu(t)$ may be hard to produce, if not impossible. In cases where $\mu(t)$ is highly unpredictable, organisms should have enough space to collect resources during times of greatest predicted scarcity, particularly if the changes in resource abundance occur rapidly, often, or for long periods of time (erratic $\mu(t)$). -->

The leftmost column of fig. \@ref(fig:5-5-sims) <!--(constant $\e(R)$) -->illustrates the effects of $\var(R)$ on the simulated space-use requirements while $\e(R)$ remains constant. Overall, both mean $H$ and the variance around it increase with $\sigma^2(t)$ (most visible with constant $\e(R)$ and linear $\var(R)$). This is because, similarly to resource-poor periods, times of greater stochasticity require the organism to move over larger areas for longer periods of time. This results in a greater uncertainty in how much time and space the organism will require to reach satiety, or indeed whether an organism living in highly stochastic environments can even reach satiety within a finite amount of time.

The remaining panels in fig. \@ref(fig:5-5-sims) illustrate $\e(R)$ and $\var(R)$ jointly affect $H$ and how complex the effects can be. Since $\e(R)$ and $\var(R)$ have opposite effects on $H$, disentangling the effects can be particularly difficult when both parameters change in a correlated manner (e.g., linear $\e(R)$ and $\var(R)$). When both $\e(R)$ and $\var(R)$ increase linearly, $H$ initially increases since the effect of $\var(R)$ is stronger, but then decreases as the effect of $\e(R)$ begins to dominate. Difficulties in disentangling the two effects are explored in greater depth in the case study in the following section.

<!-- Not all 25 scenarios depicted in fig. \@ref(fig:5-5-sims) may be realistic, but the trends in $\e(R)$ and $\var(R)$, and their impacts on space use are useful examples that can be thought of as simplified scenarios. $\e(R)$ and $\var(R)$ may be assumed to be (approximately) constant in highly homogeneous environments. Although it is impossible for $\mu(t)$ and $\sigma^2(t)$ to increase linearly forever, these examples are useful to demonstrate that linear changes in $\mu(t)$ and $\sigma^2(t)$ affect $H$ nonlinearly. Cyclical oscillations in $\e(R)$ and $\var(R)$ may occur in urban environments as human activity changes within and between days [@peron_periodic_2017; @ikeda_home_2022] and as temperatures fluctuate daily and seasonally [@geremia_migrating_2019; @alston_temperature_2020], while $\e(R)$ and $\var(R)$ may drift randomly in highly complex environments which are too hard to predict. Finally, erratic changes in $\e(R)$ and $\var(R)$ may occur in environments where changes are very sudden, such as areas prone to fires or floods, as well as habitats with drastic human alteration (e.g., a forest which is clear-cut with a subsequent artificial re-forestation). However, if stochastic changes are sufficiently small and frequent, organisms may perceive them as continuous and smooth changes rather a series of small and sudden changes. -->

Although the temporal trends in fig. \@ref(fig:5-5-sims) are complex and the effects of $\e(R)$ and $\var(R)$ can be hard to disentangle, two simple relationships emerge when $H$ is shown as a function of either $\e(R)$ or $\var(R)$, rather than time (panels A and B of fig. \@ref(fig:5-5-reg-mean)). The estimated relationships follow the hypotheses we presented in figs. \@ref(fig:mu-hyp) and \@ref(fig:s2-hyp), although we found that the effect of $\var(R)$ at average $\e(R)$ was linear with a slight sublinear saturation as $\var(R)$ increased.<!--We ensured this was not due to a high minimum or low maximum $\var(R)$ (fig. \@ref(fig:s2-hyp)) by simulating within the range $\sigma^2(t) \in [2\,500, 250\,000] \implies \sigma \in [50, 500]$, since the range of $\e(R)$ was $\mu(t) \in [100, 200]$.--> However, notice that the effect of $\var(R)$ on $E(H)$ depends strongly on $\e(R)$ (panel C). The increase in $\e(H)$ over $\var(R)$ is strongly nonlinear at low values of $\e(R)$, but it is not affected by $\var(R)$ much when $\e(R)$ is very low (and $\e(H)$ is consequently very high).

<!-- While we modeled the effects of $\e(R)$ and $\var(R)$ on both the 95% and the 50% space-use requirements, we only show results for the 95^th^ percentile because the results the 50^th^ quantile followed the same trends, other than the multiplicative rescaling required to account for the difference in magnitude between the two quantiles.  -->

```{r 5-5-reg-mean, eval=TRUE, echo=FALSE, fig.cap="Effects of $\\e(R)$ and $\\var(R)$ on simulated spatial requirements with 95\\% Bayesian credible intervals for the mean. While the effect of $\\var(R)$ is sublinear in panel B, the effect of $\\var(R)$ depends strongly on $\\e(R)$ (see panel C).\nCredible intervals were calculated using 10,000 samples from the posterior distribution. The relationships were estimated using a Generalized Additive Model for Location and Scale with a Gamma location-scale family of distributions ($\\tt{mgcv\\!::\\!gammals})$. Details on the model structure are provided in Appendix B.", out.width='100%'}
knitr::include_graphics('../figures/simulation-regression-mean.png')
```

As expected by the changes in the spread of the points in panels A and B of figure \@ref(fig:5-5-reg-mean), the variance in $H$ also depends on $\e(R)$ and $\var(R)$ (figure \@ref(fig:5-5-reg-var)A-B). Here, the effects of the two variables are more pronounced, even for low values of $\e(R)$ or high values of $\var(R)$. Consequences of these effects are explored in the discussion section.

```{r 5-5-reg-var, eval=TRUE, echo=FALSE, fig.cap="Effects of $\\e(R)$ and $\\var(R)$ on the variance in simulated spatial requirements, with 95\\% Bayesian credible intervals for the estimated variance. Credible intervals were calculated using 10,000 samples from the posterior distribution. Additional details on the model structure are provided in Appendix B.", out.width='100%'}
knitr::include_graphics('../figures/simulation-regression-var.png')
```

# A case study on a lowland tapir in the Brazilian Cerrado

\noindent The simulations in the section above mostly follow the hypotheses we presented in the introduction, but they are based on assumptions that are often not met in real natural enviroments. Organisms live in spatiotemporally heterogeneous and dynamic environments, which the development of perceptual ranges, navigation, and memory. Together, these abilities result in selective space use that depends on resource availability [@kacelnik_ideal_1992] and resource depletion [@charnov_optimal_1976].

In this section, we test the hypotheses illustrated in the introduction using empirical tracking data on a lowland tapir from the Brazilian Cerrado along with estimates of $\e(R)$ and $\var(R)$ using the Normalized Difference Vegetation Index [NDVI, see @pettorelli_normalized_2011], a remote-sensed measure of lanscape greeness, as a proxy of forage abundance. Appendix C contains additional information on how NDVI was modeled along with details on the continuous-time movement models [@fleming_ctmm_2021; @noonan_scale-insensitive_2019] and autocorrelated kernel density estimation [@noonan_comprehensive_2019; @alston_mitigating_2022; @silva_autocorrelationinformed_2022] used to quantify the tapir's space-use requirements. Through this example, we aim to provide support for the two hypotheses using empirical data that do not depend on as strong assumptions as those required by the simulations in the section above. Concurrently, we also hope to show the potential complications that arise with empirical data.

<!-- The hypotheses we present here allow researchers to combine otherwise complicated findings (e.g., fig. \@ref(fig:5-5-sims)) and generalize results to a single, common set of functions. We expect the two hypotheses we present here (figs. \@ref(fig:mu-hyp) and \@ref(fig:s2-hyp)) to be applicable to all motile organisms, once differences in size, metabolic needs, and diet are accounted for. In this subsection, we illustrate how researchers can leverage these hypotheses using their own data to build on previous work rather than listing the results as new findings. Additionally, the analyses can be expanded to the population or species level with the use of Hierarchical Generalized Additive Models (HGAMs) with individual-level intercepts and smooth terms [see models "GS" and "GI" in @pedersen_hierarchical_2019]. Such models would allow one to test whether members of the same population or species respond similarly to $\e(R)$ and $\var(R)$ while also quantifying any deviation from the hypothesized behavior. Individuals and populations that were once thought to have different space-use requirements due to differences in behavior may instead be simply responding to different local conditions along the functions in figs. \@ref(fig:mu-hyp) and \@ref(fig:s2-hyp) [e.g., @singh_migration_2012]. In the following sections, we illustrate how one can apply this framework to empirical data. -->
